### 2024-09-26 Paper/MSEPM-Manuscript-Updated.tex
Paper update

\documentclass{article}
\usepackage{latexsym}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{epsfig}
\usepackage{caption}
% \usepackage{psfig}
%\usepackage{epic}
%\usepackage{eepic}
\usepackage{colordvi}
%\usepackage{wrapfig}  
\usepackage{afterpage}
%\usepackage{placeins}
\usepackage{graphics,color}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{placeins}
\usepackage{longtable}
\usepackage{color}      % Need the color package
\usepackage{epsfig}
\usepackage{subfiles}
\usepackage[style=nature]{biblatex}
\usepackage{float}
\usepackage{lineno}

@@

\bibliography{MSEPM-Manuscript}

@@

\abstract{Epigenetic clocks, DNA methylation based predictive models of chronological age, are often utilized to study aging associated biology. 
Despite their widespread use, these methods do not account for other factors that also contribute to the variability of DNA methylation data.
 For example, many CpG sites show strong sex-specific or cell type specific patterns that likely impact the predictions of epigenetic age.  
 To overcome these limitations, we developed a multidimensional extension of the Epigenetic Pacemaker, the Multi-State Epigenetic Pacemaker (MSEPM). 
 We show that the MSEPM is capable of accurately modeling multiple methylation associated factors simultaneously, while also providing site specific 
 models that describe the per site relationship between methylation and these factors. We utilized the MSEPM with a large aggregate cohort of blood
  methylation data to construct models of the effects of age, sex and cell type heterogeneity on DNA methylation.  We found that these models 
  capture a large faction of the variability at thousands of DNA methylation sites.  Moreover, this approach allows us to identify sites that 
  are primarily affected by aging and no other factors.  An analysis of these sites reveals that those that lose methylation over time are 
  enriched for CTCF transcription factor chip peaks, while those that gain methylation over time are associated with bivalent promoters of 
  genes that are not expressed in blood. These observations suggest mechanisms that underlie age associated methylation changes and suggest 
  that age associated increases in methylation may not have strong functional consequences on cell states. In conclusion, the MSEPM is 
  capable of accurately modeling multiple methylation associated factors and the models produced can illuminate site specific 
  combinations of factors that affect methylation dynamics.}

@@

We implemented a simulation framework using the MSEPM formulation to evaluate the performance of the MSEPM model under various conditions.
To simulate the association between methylation status and an observable factor we modeled the 
epigenetic factor $p_{k,j}$ as function of time, or age, and magnitude, $p_{k,j}^0$ with a non-linear transformation $\gamma_{k}$, 
where $p_{k,j} = Age_j p_{k,j}^{0,\gamma_{k}}$. In practice the value of the $p_{k,j}$ is often unknown and the association between 
methylation status and $p_{k,j}$ is inferred through the observable factor $p_{k,j}^0$.

@@

\subsection{H3K4me3 enrichment analysis} 
Enrichment of analysis for H3K4me3 (Figure 7A) was carried out by downloading rpm normalized bigwig files of H3K4me3 ChIP-seq data from 
epigenomesportal\cite{Bujold2016-vk} for CD38+ B Cells and CD56+ NTK Cells (for both 0-5 years old and 60-65 years old individuals). 
Heatmaps of H3K4me3 were generated using deepTools2\cite{Ramirez2016-xl} using the computeMatrix and plotHeatmap function to plot the 
bigwig signal over genomic regions of cluster 7 as the BED input. The IGV genome browser\cite{Robinson2011-he} was used to generate an image of the KCTD1 
and IRS2 promoter regions shown in Figure 7B using downloaded bigwig tracks.

@@

We simulated 90 methylation sites (Figure 1I). We then evaluated the MSEPM model as follows. We simulated 1000 samples with the 
four epigenetic factors described above. We then simulated methylation values using the simulated site rates. Simulated samples
 were then split for training ($n=500$) and testing ($n=500$). MSEPM models were then fitted using the values of the input factors, $p^0_{k,j}$. 
 We generated 1000 simulated datasets and fit MSEPM models using four combinations of input factors (Age, Age-Sex, Age-Sex-CN, Age-Sex-CN-CE). 
 Within each simulation, epigenetic state predictions and methylation site predictions were made for all testing samples. All models captured 
 the nonlinear association between simulated age and methylation (Supp. Figure 1). As the number of factors in the model is increased the mean 
 absolute error (MAE) between the predicted epigenetic states and the simulated epigenetic factors decreases (Figure 2A). Importantly, to accurately
  assess simulated age it is necessary to account for the influence of the other simulated factors (Sex, CN, CE).

@@

The site clusters largely conform to underlying biological expectations. Cluster one contains sites 
that are wholly associated with sex status and localized to the X chromosome (Supp. Table 1) and is enriched for peaks of 
transcription factors associated with sex specific regulation such as MAZ\cite{Lopes-Ramos2020-ex}. 
Clusters nine and ten contain sites whose methylation status is largely driven by CT-PC1, and are enriched for 
transcription factors associated with granulocyte development (CEBPB, CEBPA, EP300, ETV6)\cite{Theilgaard-Monch2022-zw,Guerzoni2006-ii}. 
Similarly, clusters two, five and eight are associated with CT-PC2 and are enriched for 
transcription factor peaks associated with immune development (ZBED1, ETV6, FOSL2, FOS, TBX21). Clusters four and 
six are associated with loss of methylation with age. Cluster six is highly enriched for CTCF binding sites; CTCF 
is known to increase at sites where methylation is lost during aging\cite{Tharakan2020-pj}. Cluster four is enriched 
for STAT3 whose activation during exercise is age dependent\cite{Mohamed2020-he,Trenerry2008-kj}.  
Cluster seven is associated with the accumulation of methylation with age and is enriched for immunoprecipitation peaks
 for aging associated transcription factors SMAD4 and RE1-Silencing Transcription Factor (REST). SMAD4 encodes a protein 
 involved in the transforming growth factor beta (TGF-$\beta$) signaling pathway. Age related dysregulation of TGF-$\beta$ 
 has been linked to reduced skeletal muscle regeneration\cite{Carlson2009-uz,Paris2016-fo} and SMAD4 polymorphisms 
 are associated with longevity\cite{Carrieri2004-by}. REST is a transcriptional repressor of neuron specific genes in 
 non-neuronal cells\cite{Chong1995-dj,Coulson2005-pb}. REST expression is upregulated in aged prefrontal cortex 
 tissue and the absence of REST expression is associated with cognitive impairment\cite{Lu2014-dz} and 
 cellular senescence in neurons\cite{Rocchi2021-od}.

@@

Based on these results we hypothesize that the mechanisms that underlie the gain of methylation with age at these 
bivalent promoters is the age-associated loss of H3K4me3.  It is well established that the presence of trimethylation
 on H3K4 inhibits de novo methylation, and this effect explains the hypomethylation that is typical of promoters, 
 including bivalent promoters.  We therefore hypothesize that the gain of methylation at these sites may be caused 
 by an age associated loss of H3K4me3.  In order to demonstrate that H3K4me3 decreases with age for genomic regions 
 where DNA methylation increases, we used published H3K4me3 ChIP-seq data from epigenomesportal\cite{Bujold2016-vk}. 
 We selected two different blood cell types CD38+ B Cells and CD56+ NTK Cells and plotted the H3K4me3 signal of young 
 (0 to 5 years old) versus old individuals (60 to 65 years old) over genomic regions of cluster 7 (Figure 7A). 
 Our analysis shows that younger individuals have higher levels of H3K4me3 compared to older ones (Figure 7A) as 
 also shown for two selected genomic loci of cluster 7 (the promoters of KCTD1 and IRS2 genes) where we can observe
 a marked decrease in the levels of H3K4me3 as age increases (Figure 7B). All together these data suggest that genomic 
 regions whose DNA methylation is increased with age exhibit an age dependent loss of H3K4me3, thus showing an inverse 
 correlation between DNA methylation and H3K4me3 at these genomic loci.

@@

In conclusion, we introduced a multi-dimensional extension of the Epigenetic Pacemaker, the MSEPM. The MSEPM is capable of 
accurately modeling multiple methylation associated factors simultaneously. This paradigm can elucidate the site specific 
regulation underpinning methylome dynamics. It allows us to characterize the mechansims underlying age assocated increases 
in methylation sites, suggesting that these were caused by the loss of H3K4Me3 at bivalent promoters of genes that are silenced 
in blood.  The MSEPM is available under the MIT license at https://github.com/NuttyLogic/MultistateEpigeneticPacemaker.

@@

\printbibliography

@@

\begin{center}
    \begin{figure}
    \includegraphics[scale=.25]{Figures/Figure1.png}    
    \footnotesize
    \caption*{\small \textbf{Figure 1:}Simulated factors and the association with simulated methylation values. (A) Age with a non-linear association with methylation 
(B). Sex (C) with a binary association with methylation (D). Normal factor (E) with a linear relationship with methylation (F). Continuous exponential
 trait (G) with a linear relationship with methylation. (I) Simulated methylation sites. Each simulation site has a starting methylation value $r^0_i$ ,
  rate of change associated with each simulated factor $r_{i, factor}$ and range of variation $\delta_i$.}
    \end{figure}
\end{center}

\begin{center}
    \begin{figure}
    \includegraphics[scale=.2]{Figures/Figure2.png}
    \footnotesize
    \caption*{\small \textbf{Figure 2:} (A) The MAE of the factor predictions on the testing set as multiple factors are modeled simultaneously and 
    (B) predicted methylation MAE as factors are included in the MSEPM model where the centerline is the 50th quantile and the box 
    with greatest width contains 50\% of the underlying data with each smaller box containing 50\% of the remaining data 
    with 6 levels of box width. (C) Model coefficients for 
Age, Sex, Continuous Normal and Continuous Exponential factors for models trained $(n = 500)$ with all four simulated 
factors. (D). Simulated and predicted methylation values for all simulated testing sites across all training fold }
    \end{figure}
\end{center}

\begin{center}
    \begin{figure}
    \includegraphics[scale=.2]{Figures/Figure3.png}
    \footnotesize
    \caption*{\small \textbf{Figure 3:} Distribution of age (A) and (B) sex in aggregate blood dataset. (C) Calculated cell type composition and (D) loading plot of
 principal components of cell type composition in aggregate blood data set.}
    \end{figure}
\end{center}

\begin{raggedleft}
\begin{figure}
\includegraphics[scale=.15]{Figures/Figure4.png}
\footnotesize
\caption*{\small \textbf{Figure 4:} MSEPM model trained with age, sex, CT-PC1 and CT-PC2 predictions within testing 
set for epigenetic factors (A) age, (B) sex, (C) CT-PC1 and (D) CT-PC2. (E) Observed and predicted
 methylation values for training set has high concordance $(R^2=0.933, MAE=0.035)$}
\end{figure}
\end{raggedleft}

\begin{center}
    \begin{figure}
    \includegraphics[scale=.15]{Figures/Figure5.png}
    \footnotesize
    \caption*{\small \textbf{Figure5: } (A) Site clustering by standardized model coefficients. Sites clusters show 
    distinct relationships with modeled traits. (B-K) Top five enriched transcription factors for clusters 1 - 10. 
    }
    \end{figure}
\end{center}

\begin{center}
    \begin{figure}
    \includegraphics[scale=.15]{Figures/Figure6.png}
    \footnotesize
    \caption*{\small \textbf{Figure6: } (A) Heatmap of H3K4me3 ChIP-seq enrichment for two different blood cell types (CD38\textsuperscript{+} B Cells and CD56\textsuperscript{+} NTK Cells) in 
two cohorts of individual within 0 to 5 years old and 60 to 65 years old. The average level within 2kb up and downstream for centered genomic regions of cluster 7 is 
represented above the heatmap. (B) Genome browser view of H3K4me3 levels in each cohort at the promoter regions of \textit{KCTD1} and \textit{IRS2} genes.  
    }
    \end{figure}
\end{center}

\begin{center}
    \begin{figure}
    \includegraphics[scale=.15]{Figures/Figure7.png}
    \footnotesize
    \caption*{\small \textbf{Figure7: } (A) Heatmap of H3K4me3 ChIP-seq enrichment for two different blood cell types (CD38\textsuperscript{+} B Cells and CD56\textsuperscript{+} NTK Cells) in 
two cohorts of individual within 0 to 5 years old and 60 to 65 years old. The average level within 2kb up and downstream for centered genomic regions of cluster 7 is 
represented above the heatmap. (B) Genome browser view of H3K4me3 levels in each cohort at the promoter regions of \textit{KCTD1} and \textit{IRS2} genes. }
    \end{figure}
\end{center}

>>>>

\documentclass[sn-nature]{sn-jnl}
\usepackage{latexsym}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{epsfig}
\usepackage{caption}
% \usepackage{psfig}
%\usepackage{epic}
%\usepackage{eepic}
\usepackage{colordvi}
%\usepackage{wrapfig}  
\usepackage{afterpage}
%\usepackage{placeins}
\usepackage{graphics,color}
\usepackage{graphicx}
\usepackage{placeins}
\usepackage{longtable}
\usepackage{color}      % Need the color package
\usepackage{epsfig}
\usepackage{subfiles}
\usepackage{float}
\usepackage{lineno}

@@

@@

\section{Abstract}\label{abstract}

Epigenetic clocks, DNA methylation based predictive models of chronological age, are often utilized to study aging associated biology. 
Despite their widespread use, these methods do not account for other factors that also contribute to the variability of DNA methylation data.
 For example, many CpG sites show strong sex-specific or cell type specific patterns that likely impact the predictions of epigenetic age.  
 To overcome these limitations, we developed a multidimensional extension of the Epigenetic Pacemaker, the Multi-State Epigenetic Pacemaker (MSEPM). 
 We show that the MSEPM is capable of accurately modeling multiple methylation associated factors simultaneously, while also providing site specific 
 models that describe the per site relationship between methylation and these factors. We utilized the MSEPM with a large aggregate cohort of blood
  methylation data to construct models of the effects of age, sex and cell type heterogeneity on DNA methylation.  We found that these models 
  capture a large faction of the variability at thousands of DNA methylation sites.  Moreover, this approach allows us to identify sites that 
  are primarily affected by aging and no other factors.  An analysis of these sites reveals that those that lose methylation over time are 
  enriched for CTCF transcription factor chip peaks, while those that gain methylation over time are associated with bivalent promoters of 
  genes that are not expressed in blood. These observation suggest mechanisms that underlie age associated methylation changes and suggest 
  that age associated increases in methylation may not have strong functional consequences on cell states. In conclusion, the MSEPM is 
  capable of accurately modeling multiple methylation associated factors and the models produced can illuminate site specific 
  combinations of factors that affect methylation dynamics.

@@

We implemented a simulation framework using the MSEPM formulation evaluate the performance of the MSEPM model under various conditions.
To simulate the association between methylation status and an observable factor we modeled the 
epigenetic factor $p_{k,j}$ as function of time, or age, and magnitude, $p_{k,j}^0$ with a non-linear transformation $\gamma_{k}$, 
where $p_{k,j} = Age_j p_{k,j}^{0,\gamma_{k}}$. In practice the value of the $p_{k,j}$ is often unknown and the association between 
methylation status and $p_{k,j}$ is inferred through the observable factor $p_{k,j}^0$.

@@

\subsection{H3K4me3 enrichment analysis} 
Enrichment of analysis for H3K4me3 (figure 7A) was carried out by downloading rpm normalized bigwig files of H3K4me3 ChIP-seq data from 
epigenomesportal\cite{Bujold2016-vk} for CD38+ B Cells and CD56+ NTK Cells (for both 0-5 years old and 60-65 years old individuals). 
Heatmaps of H3K4me3 were generated using deepTools2\cite{Ramirez2016-xl} using the computeMatrix and plotHeatmap function to plot the 
bigwig signal over genomic regions of cluster 7 as the BED input. The IGV genome browser\cite{Robinson2011-he} was used to generate an image of the KCTD1 
and IRS2 promoter regions shown in figure 7B using downloaded bigwig tracks.

@@

We simulated 90 methylation sites (Figure 1I). We then evaluated the MSEPM model as follows. We simulated 1000 samples with the 
four epigenetic factors described above. We then simulated methylation values using the simulated site rates. Simulated samples
 were then split for training ($n=500$) and testing ($n=500$). MSEPM models were then fitted using the values of the input factors, $p^0_{k,j}$. 
 We generated 1000 simulated datasets and fit MSEPM models using four combinations of input factors (Age, Age-Sex, Age-Sex-CN, Age-Sex-CN-CE). 
 Within each simulation, epigenetic state predictions and methylation site predictions were made for all testing samples. All models captured 
 the nonlinear association between simulated age and methylation (Supp. Figure 1). As the number of factors in the model is increased the mean 
 absolute error (MAE) between the predicted epigenetic states and the simulated epigenetic factors decreases (Figure 2A). Importantly, to accurately
  assess simulated age it is necessary to account for the influence of the other simulated factors (Sex, CN, CE). The MAE between the predicted 
  and simulated methylation values decreases as simulated factors are added to the model, and accurately assessing the methylation status of a 
  simulated site requires that the factor associated with the methylation status at the site is included in the model (Figure 2A).

@@

The site clusters largely conform to underlying biological expectations. Cluster one contains sites 
that are wholly associated with sex status and localized to the X chromosome (Supp. Table 1) and is enriched for peaks of 
transcription factors associated with sex specific regulation such as MAZ\cite{Lopes-Ramos2020-ex}. 
Clusters nine and ten contain sites whose methylation status is largely driven by CT-PC1, and are enriched for 
transcription factors associated with granulocyte development (CEBPB, CEBPA, EP300, ETV6)\cite{Theilgaard-Monch2022-zw,Guerzoni2006-ii}. 
Similarly, clusters two, five and eight are associated with CT-PC2 and are enriched for 
transcription factor peaks associated with immune development (ZBED1, ETV6, FOSL2, FOS, TBX21). Clusters four and 
six are associated with loss of methylation with age. Cluster six is highly enriched for CTCF binding sites; CTCF 
is known to increase at sites where methylation is lost during aging\cite{Tharakan2020-pj}. Cluster four is enriched 
for STAT3 whose activation during exercise is age dependent\cite{Mohamed2020-he,Trenerry2008-kj}.  
Cluster seven is associated with the accumulation of methylation with age and is enriched for immunoprecipitation peaks
 for aging associated transcription factors SMAD4 and RE1-Silencing Transcription Factor (REST). SMAD4 encodes a protein 
 involved in the transforming growth factor beta (TGF-$\beta$) signaling pathway. Age related dysregulation of TGF-$\beta$ 
 has been linked to reduced skeletal muscle regeneration\cite{Carlson2009-uz,Paris2016-fo} and SMAD4 polymorphisms 
 are associated with longevity\cite{Carrieri2004-by}). REST is a transcriptional repressor of neuron specific genes in 
 non-neuronal cells\cite{Chong1995-dj,Coulson2005-pb}. REST expression is upregulated in aged prefrontal cortex 
 tissue and the absence of REST expression is associated with cognitive impairment\cite{Lu2014-dz} and 
 cellular senescence in neurons\cite{Rocchi2021-od}).

@@

Based on these results we hypothesize that the mechanisms that underlies the gain of methylation with age at these 
bivalent promoters is the age-associated loss of H3K4me3.  It is well established that the presence of trimethylation
 on H3K4 inhibits de novo methylation, and this effect explains the hypomethylation that is typical of promoters, 
 including bivalent promoters.  We therefore hypothesize that the gain of methylation at these sites may be caused 
 by an age associated loss of H3K4me3.  In order to demonstrate that H3K4me3 decreases with age for genomic regions 
 where DNA methylation increases, we used published H3K4me3 ChIP-seq data from epigenomesportal\cite{Bujold2016-vk}. 
 We selected two different blood cell types CD38+ B Cells and CD56+ NTK Cells and plotted the H3K4me3 signal of young 
 (0 to 5 years old) versus old individuals (60 to 65 years old) over genomic regions of cluster 7 (Figure 7A). 
 Our analysis shows that younger individuals have higher levels of H3K4me3 compared to older ones (Figure 7A) as 
 also shown for two selected genomic loci of cluster 7 (the promoters of KCTD1 and IRS2 genes) where we can observe
 a marked decrease in the levels of H3K4me3 as age increases (Figure 7B). All together these data suggest that genomic 
 regions whose DNA methylation is increased with age exhibit an age dependent loss of H3K4me3, thus showing an inverse 
 correlation between DNA methylation and H3K4me3 at these genomic loci.

@@

In conclusion, we introduced a multi-dimensional extension of the Epigenetic Pacemaker, the MSEPM. The MSEPM is capable of 
accurately modeling multiple methylation associated factors simultaneously. This paradigm can elucidate the site specific 
regulation underpinning methylome dynamics. It allows us to characterize the mechansims underlying age assocated increases 
in methylation sites, suggeting that these were caused by the loss of H3K4Me3 at bivalent promoters of genes that are silenced 
in blood.  The MSEPM is available under the MIT license at https://github.com/NuttyLogic/MultistateEpigeneticPacemaker.

@@

\bibliography{MSEPM-Manuscript-Updated.bib}

@@

\begin{center}
    \begin{figure}
    \includegraphics[scale=.25]{Figure1.png}    
    \footnotesize
    \caption*{\small \textbf{Figure 1:}Simulated factors and the association with simulated methylation values. (A) Age with a non-linear association with methylation 
(B). Sex (C) with a binary association with methylation (D). Normal factor (E) with a linear relationship with methylation (F). Continuous exponential
 trait (G) with a linear relationship with methylation. (I) Simulated methylation sites. Each simulation site has a starting methylation value $r^0_i$ ,
  rate of change associated with each simulated factor $r_{i, factor}$ and range of variation $\delta_i$.}
    \end{figure}
\end{center}

\begin{center}
    \begin{figure}
    \includegraphics[scale=.2]{Figure2.png}
    \footnotesize
    \caption*{\small \textbf{Figure 2:} (A) The MAE of the factor predictions on the testing set as multiple factors are modeled simultaneously and 
    (B) predicted methylation MAE as factors are included in the MSEPM model where the centerline is the 50th quantile and the box 
    with greatest width contains 50\% of the underlying data with each smaller box containing 50\% of the remaining data 
    with 6 levels of box width. (C) Model coefficients for 
Age, Sex, Continuous Normal and Continuous Exponential factors for models trained $(n = 500)$ with all four simulated 
factors. (D). Simulated and predicted methylation values for all simulated testing sites across all training fold }
    \end{figure}
\end{center}

\begin{center}
    \begin{figure}
    \includegraphics[scale=.2]{Figure3.png}
    \footnotesize
    \caption*{\small \textbf{Figure 3:} Distribution of age (A) and (B) sex in aggregate blood dataset. (C) Calculated cell type composition and (D) loading plot of
 principal components of cell type composition in aggregate blood data set.}
    \end{figure}
\end{center}

\begin{raggedleft}
\begin{figure}
\includegraphics[scale=.15]{Figure4.png}
\footnotesize
\caption*{\small \textbf{Figure 4:} MSEPM model trained with age, sex, CT-PC1 and CT-PC2 predictions within testing 
set for epigenetic factors (A) age, (B) sex, (C) CT-PC1 and (D) CT-PC2. (E) Observed and predicted
 methylation values for training set has high concordance $(R^2=0.933, MAE=0.035)$}
\end{figure}
\end{raggedleft}

\begin{center}
    \begin{figure}
    \includegraphics[scale=.15]{Figure5.png}
    \footnotesize
    \caption*{\small \textbf{Figure5: } (A) Site clustering by standardized model coefficients. Sites clusters show 
    distinct relationships with modeled traits. (B-K) Top five enriched transcription factors for clusters 1 - 10. 
    }
    \end{figure}
\end{center}

\begin{center}
    \begin{figure}
    \includegraphics[scale=.15]{Figure6.png}
    \footnotesize
    \caption*{\small \textbf{Figure6: } (A) Heatmap of H3K4me3 ChIP-seq enrichment for two different blood cell types (CD38\textsuperscript{+} B Cells and CD56\textsuperscript{+} NTK Cells) in 
two cohorts of individual within 0 to 5 years old and 60 to 65 years old. The average level within 2kb up and downstream for centered genomic regions of cluster 7 is 
represented above the heatmap. (B) Genome browser view of H3K4me3 levels in each cohort at the promoter regions of \textit{KCTD1} and \textit{IRS2} genes.  
    }
    \end{figure}
\end{center}

\begin{center}
    \begin{figure}
    \includegraphics[scale=.15]{Figure7.png}
    \footnotesize
    \caption*{\small \textbf{Figure7: } (A) Heatmap of H3K4me3 ChIP-seq enrichment for two different blood cell types (CD38\textsuperscript{+} B Cells and CD56\textsuperscript{+} NTK Cells) in 
two cohorts of individual within 0 to 5 years old and 60 to 65 years old. The average level within 2kb up and downstream for centered genomic regions of cluster 7 is 
represented above the heatmap. (B) Genome browser view of H3K4me3 levels in each cohort at the promoter regions of \textit{KCTD1} and \textit{IRS2} genes. }
    \end{figure}
\end{center}
